---
title: "Data Prepration"
execute:
  warning: false
---
# 1.Install Package
```{r}
pacman::p_load(sf, tmap, tidyverse, ggplot2, plotly, dplyr,readr, data.table)
```

# 2.Data Prepration
## 2.1 Data Prepration 
  We aggregate monthly observations from 2013 to 2023, sourced from four distinct weather stations strategically located across Singapore's geographic regions. This process involves the meticulous integration of data sets to represent a comprehensive view of Singapore's climatic conditions over the specified period.
  This consolidated data set serves as a foundation for futher climate analysis, trend identification, and regional weather comparisons within Singapore.
```{r}
#| code-fold: true
# We will import all the monthly data from 2013 to 2023 chosen from five different region.
setwd("D:/OtisChou/ISSS608-G1-Group10-Project/Actualwork/Dataprep/data")
files <- list.files(pattern = "*.csv", full.names = TRUE)
weatherdata <- rbindlist(lapply(files, fread, encoding = "UTF-8"))

# Delete Unwanted columns
weatherdata <- weatherdata %>% select(-c("Highest 60 Min Rainfall (mm)", "Highest 30 Min Rainfall (mm)","Highest 120 Min Rainfall (mm)"))

# Change the Name of the column
new_column_names <- c("Station", "Year", "Month","Day","Rainfall","MeanTemp","MaxTemp","MinTemp","MeanWind","MaxWind")
names(weatherdata) <- new_column_names

#replace \97x with NA
weatherdata <- weatherdata %>%
  mutate(across(where(is.character), ~na_if(., '\x97')))

#Convert data type
weatherdata <- weatherdata %>%
  mutate(across(c(Rainfall,MeanTemp,MaxTemp,MinTemp,MeanWind,MaxWind), as.numeric))%>%
  mutate(across(c(Year,Month,Day), as.character))

```
## 2.2 Saved Consolidated Data For Future Use
In this phase, we create a rds format file to save our consolidated data set.
```{r}
write_rds(weatherdata,
          "data/weatherdata.rds")
```

## 2.3 Read Output Data
In this phase, we put our consolidated data set into the R programming environment, ensuring a seamless integration for comprehensive evaluation. 
```{r}
weatherdata <- read_rds("data/weatherdata.rds")
summary(weatherdata)
```